\bf{Comparability of child noncognitive skills across UK birth cohorts}

### 2018-02-28 codes/R/slides/cohorts_nc_slides.Rnw
slides discussed 28/2/18

\begin{column}{0.6\textwidth}
\bi
\item Data from 5 year follow-up, maternal questionnaire
\item Rutter A scale \citep{Rutter1970}
  \bi
  \item 19 items (behaviours)
  \item Options: \textit{Doesn't apply, Applies somewhat, Certainly applies}
  \item Usually conceptualised as measuring two dimensions
  \bi \item \emph{externalising} behaviour problems (anti-social) \item \emph{internalising} behaviour problems (neurotic) \ei
  \ei
  \item Additional questions from maternal module
\ei
\end{column}

@@

\bi
\item Data from 5 year follow-up (MCS3), maternal questionnaire
\item Complex survey design: need to rebalance data to make representative
\item Strengths and Difficulties Questionnaire (SDQ, \citealp{Goodman1994})
  \bi
  \item ``Updated" and expanded version of the Rutter scale 
  \item 25 items (behaviours)
  \item Options: \textit{Not true, Somewhat true, Certainly true}
  \item Designed with 5 subscales
  \item Broader subscales are sometimes justifiable \citep{Goodman2010}
  \ei
\ei

@@

\begin{center}
<<EFATAB, dependson='MCSEFA', results='asis'>>=
# combine results
efatab <- cbind(efatab.bcs,efatab.mcs)
colnames(efatab) <- c("BCS", "MCS")
rownames(efatab) <- c("Optimal Coordinates", "Acceleration Factor", "Parallel Analysis", "Kaiser", "VSS Compl. 1", "VSS Compl. 2")

print(xtable(efatab,
                                  # item num        Factor          Num cats        BCS text        BCS num           BCS Certain     BCS Somewhat    BCS Applies             MCS text          MCS num         
                    align="L{.4\\textwidth}C{.2\\textwidth}C{.2\\textwidth}",
                    digits = 0
                    ), 
             size="\\small",
             include.rownames=T,
             include.colnames=T,
             booktabs = T,
             sanitize.text.function = function(x){x},
             tabular.environment = "tabular", width = ".8\\textwidth",
             floating = FALSE
)
@
\end{center}
\end{frame}

@@

<<FA_INVARIANCE, dependson='LOAD_DATA'>>=
# estimate models of different levels of invariance
@

@@

% # <<FA_SCORES_COMPARE, dependson='FA_SCORES'>>=
% # # assemble and compare scores across different models
% # @

<<FA_SCORES_SELECT, dependson='FA_FINAL'>>=
# select model to use for inequality analysis
@

@@

\begin{frame}
\frametitle{Factor scores vs. Sum scores}
\begin{figure}
<<FACRAW, dev='tikz', fig.height=7, fig.width=7, out.width='.65\\linewidth'>>=
@
\end{figure}
\end{frame}

@@

<<LOESS_REAL, dependson="FA_SCORES_SELECT">>=
@

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Visualising the gradient -- inflation- and growth-adjusted income}
\begin{figure}
<<LOESS_REAL_EXT, dev='tikz', fig.height=8, fig.width=12, out.width='.95\\linewidth'>>=
ploess.real.ext
@
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Visualising the gradient -- inflation- and growth-adjusted income}
\begin{figure}
<<LOESS_REAL_INT, dev='tikz', fig.height=8, fig.width=12, out.width='.95\\linewidth'>>=
ploess.real.int
@
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Quantifying the gradient -- inflation- and growth-adjusted income}
\centering
\resizebox{!}{4cm}{%
<<FA_REGS_REAL, dependson='FA_SCORES', results='asis'>>=
# regress scores on income
library(dplyr)
library(stargazer)

@@

stargazer(r.ext.m.q, r.int.m.q, r.ext.f.q, r.int.f.q,
          column.labels=c("Males","Females"), column.separate=c(2,2),
          dep.var.caption = "", multicolumn = FALSE,
          omit=c("Constant"),
          covariate.labels=c(
                paste0("Income Q2 [",toString(minmax.real[2,1]),"-",toString(minmax.real[2,2])," £/w]"),
                paste0("Income Q3 [",toString(minmax.real[3,1]),"-",toString(minmax.real[3,2])," £/w]"),
                paste0("Income Q4 [",toString(minmax.real[4,1]),"-",toString(minmax.real[4,2])," £/w]"),
                paste0("Income Q5 [",toString(minmax.real[5,1]),"-",toString(minmax.real[5,2])," £/w]"),
                             "MCS dummy",
                             "Income Q2 x MCS",
                             "Income Q3 x MCS",
                             "Income Q4 x MCS",
                             "Income Q5 x MCS"
                             ),
          no.space=TRUE, 
          omit.stat= c("rsq", "f", "ser", "aic", "ll"),
          float=FALSE    # needs begin table outside it
)
@
}
\end{frame}

@@

\begin{frame}
\frametitle{Returns}
\framesubtitle{Education returns to behaviour skills in BCS}
\centering
\resizebox{!}{4cm}{%
<<FA_RETURNS, dependson='FA_SCORES', results='asis'>>=
# regress scores on income
library(dplyr)
library(stargazer)

@@

@@

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Visualising the gradient -- inflation-adjusted income}
\begin{figure}
<<LOESS_INFL_EXT, dev='tikz', fig.height=8, fig.width=12, out.width='.95\\linewidth'>>=
ploess.infl.ext
@
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Visualising the gradient -- inflation-adjusted income}
\begin{figure}
<<LOESS_INFL_INT, dev='tikz', fig.height=8, fig.width=12, out.width='.95\\linewidth'>>=
ploess.infl.int
@
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Quantifying the gradient -- inflation-adjusted income}
\centering
\resizebox{!}{4cm}{%
<<FA_REGS_INFL, dependson='FA_SCORES', results='asis'>>=
# regress scores on income
library(dplyr)
library(stargazer)

>>>>

\begin{column}{0.6\textwidth}
\bi
\item Data from 5 year follow-up (England only), maternal questionnaire
\item Rutter A scale \citep{Rutter1970}
  \bi
  \item 19 items (behaviours)
  \item Options: \textit{Doesn't apply, Applies somewhat, Certainly applies}
  \item Usually conceptualised as measuring two dimensions
  \bi \item \emph{externalising} behaviour problems (anti-social) \item \emph{internalising} behaviour problems (neurotic) \ei
  \ei
  \item Additional questions from maternal module
\ei
\end{column}

@@

\bi
\item Data from 5 year follow-up (England only), maternal questionnaire
\item Complex survey design: rebalanced data to make representative
\item Strengths and Difficulties Questionnaire (SDQ, \citealp{Goodman1994})
  \bi
  \item ``Updated" and expanded version of the Rutter scale 
  \item 25 items (behaviours)
  \item Options: \textit{Not true, Somewhat true, Certainly true}
  \item Designed with 5 subscales
  \item Broader subscales are sometimes justifiable \citep{Goodman2010}
  \ei
\ei

@@

\begin{center}
<<EFATAB, dependson='MCSEFA', results='asis'>>=
# combine results
efatab <- cbind(efatab.bcs,efatab.mcs)
colnames(efatab) <- c("BCS All", "BCS M", "BCS F", "MCS All", "MCS M", "MCS F")

print(xtable(efatab,
                    align="L{.25\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}C{.08\\textwidth}",
                    digits = 0
                    ), 
             size="\\small",
             include.rownames=T,
             include.colnames=T,
             booktabs = T,
             sanitize.text.function = function(x){x},
             tabular.environment = "tabular", width = ".8\\textwidth",
             floating = FALSE
)
@
\end{center}
\end{frame}

@@

<<FA_INVARIANCE>>=
# estimate models of different levels of invariance
@

@@

<<FA_SCORES_MERGE, dependson='LOAD_DATA'>>=
# merge scores with other data
@

<<FA_SCORES_COMPARE, dependson='FA_SCORES'>>=
# assemble and compare scores across different models
@

<<FA_SCORES_SELECT, dependson='FA_SCORES_COMPARE'>>=
# select model to use for inequality analysis
@

@@

@@

<<LOESS_REAL, dependson='FA_SCORES_SELECT'>>=
@

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Visualising the gradient -- inflation- and growth-adjusted income}
\begin{figure}
<<LOESS_REAL_EXT, dependson='LOESS_REAL', dev='tikz', fig.height=8, fig.width=12, out.width='.95\\linewidth'>>=
ploess.real.ext
@
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Visualising the gradient -- inflation- and growth-adjusted income}
\begin{figure}
<<LOESS_REAL_INT, dependson='LOESS_REAL', dev='tikz', fig.height=8, fig.width=12, out.width='.95\\linewidth'>>=
ploess.real.int
@
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Quantifying the gradient -- inflation- and growth-adjusted income}
\centering
\resizebox{!}{4cm}{%
<<FA_REGS_REAL, dependson='FA_SCORES_SELECT', results='asis'>>=
# regress scores on income
library(dplyr)
library(stargazer)

@@

stargazer(r.ext.m.q, r.int.m.q, r.ext.f.q, r.int.f.q,
          column.labels=c("Males","Females"), column.separate=c(2,2),
          dep.var.caption = "", multicolumn = FALSE,
          omit=c("Constant"),
          covariate.labels=c(
                paste0("Income Q2 [",toString(minmax.real[2,1]),"-",toString(minmax.real[2,2])," £/w]"),
                paste0("Income Q3 [",toString(minmax.real[3,1]),"-",toString(minmax.real[3,2])," £/w]"),
                paste0("Income Q4 [",toString(minmax.real[4,1]),"-",toString(minmax.real[4,2])," £/w]"),
                paste0("Income Q5 [",toString(minmax.real[5,1]),"+ £/w]"),
                             "MCS dummy",
                             "Income Q2 x MCS",
                             "Income Q3 x MCS",
                             "Income Q4 x MCS",
                             "Income Q5 x MCS"
                             ),
          no.space=TRUE, 
          omit.stat= c("rsq", "f", "ser", "aic", "ll"),
          float=FALSE    # needs begin table outside it
)
@
}
\end{frame}

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Quantifying the gradient -- males}
\centering
\resizebox{12cm}{4cm}{%
<<FA_REGS_REAL_M, dependson='FA_SCORES', results='asis'>>=
# regress scores on income
library(dplyr)
library(stargazer)

r.ext.m.bcs <- lm(EXT ~ as.factor(incq), data=subset(scores2plot, sex=="M" & cohort=="BCS"))
r.ext2.m.bcs <- lm(EXT ~ as.factor(incq) + bwt + as.factor(smkpr) + ysch_fath + ysch_moth, data=subset(scores2plot, sex=="M" & cohort=="BCS"))
r.ext.m.mcs <- lm(EXT ~ as.factor(incq), data=subset(scores2plot, sex=="M" & cohort=="MCS"))
r.ext2.m.mcs <- lm(EXT ~ as.factor(incq) + bwt + as.factor(smkpr) + ysch_fath + ysch_moth, data=subset(scores2plot, sex=="M" & cohort=="MCS"))
r.int.m.bcs <- lm(INT ~ as.factor(incq), data=subset(scores2plot, sex=="M" & cohort=="BCS"))
r.int2.m.bcs <- lm(INT ~ as.factor(incq) + bwt + as.factor(smkpr) + ysch_fath + ysch_moth, data=subset(scores2plot, sex=="M" & cohort=="BCS"))
r.int.m.mcs <- lm(INT ~ as.factor(incq), data=subset(scores2plot, sex=="M" & cohort=="MCS"))
r.int2.m.mcs <- lm(INT ~ as.factor(incq) + bwt + as.factor(smkpr) + ysch_fath + ysch_moth, data=subset(scores2plot, sex=="M" & cohort=="MCS"))

stargazer(r.ext.m.bcs, r.ext2.m.bcs, r.int.m.bcs, r.int2.m.bcs, r.ext.m.mcs, r.ext2.m.mcs, r.int.m.mcs, r.int2.m.mcs,
          column.labels=c("BCS","MCS"), column.separate=c(4,4),
          dep.var.caption = "", multicolumn = FALSE,
          omit=c("Constant"),
          covariate.labels=c(
                             "Income Q2",
                             "Income Q3",
                             "Income Q4",
                             "Income Q5",
                             "Birthweight",
                             "Mat. Smoking in preg.",
                             "Father years of schooling",
                             "Mother years of schooling"
                             ),
          no.space=TRUE, 
          omit.stat= c("rsq", "f", "ser", "aic", "ll"),
          float=FALSE    # needs begin table outside it
)
@
}
\end{frame}

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Quantifying the gradient -- females}
\centering
\resizebox{12cm}{4cm}{%
<<FA_REGS_REAL_F, dependson='FA_SCORES', results='asis'>>=
# regress scores on income
library(dplyr)
library(stargazer)

r.ext.f.bcs <- lm(EXT ~ as.factor(incq), data=subset(scores2plot, sex=="F" & cohort=="BCS"))
r.ext2.f.bcs <- lm(EXT ~ as.factor(incq) + bwt + as.factor(smkpr) + ysch_fath + ysch_moth, data=subset(scores2plot, sex=="F" & cohort=="BCS"))
r.ext.f.mcs <- lm(EXT ~ as.factor(incq), data=subset(scores2plot, sex=="F" & cohort=="MCS"))
r.ext2.f.mcs <- lm(EXT ~ as.factor(incq) + bwt + as.factor(smkpr) + ysch_fath + ysch_moth, data=subset(scores2plot, sex=="F" & cohort=="MCS"))
r.int.f.bcs <- lm(INT ~ as.factor(incq), data=subset(scores2plot, sex=="F" & cohort=="BCS"))
r.int2.f.bcs <- lm(INT ~ as.factor(incq) + bwt + as.factor(smkpr) + ysch_fath + ysch_moth, data=subset(scores2plot, sex=="F" & cohort=="BCS"))
r.int.f.mcs <- lm(INT ~ as.factor(incq), data=subset(scores2plot, sex=="F" & cohort=="MCS"))
r.int2.f.mcs <- lm(INT ~ as.factor(incq) + bwt + as.factor(smkpr) + ysch_fath + ysch_moth, data=subset(scores2plot, sex=="F" & cohort=="MCS"))

stargazer(r.ext.f.bcs, r.ext2.f.bcs, r.int.f.bcs, r.int2.f.bcs, r.ext.f.mcs, r.ext2.f.mcs, r.int.f.mcs, r.int2.f.mcs,
          column.labels=c("BCS","MCS"), column.separate=c(4,4),
          dep.var.caption = "", multicolumn = FALSE,
          omit=c("Constant"),
          covariate.labels=c(
                             "Income Q2",
                             "Income Q3",
                             "Income Q4",
                             "Income Q5",
                             "Birthweight",
                             "Mat. Smoking in preg.",
                             "Father years of schooling",
                             "Mother years of schooling"
                             ),
          no.space=TRUE, 
          omit.stat= c("rsq", "f", "ser", "aic", "ll"),
          float=FALSE    # needs begin table outside it
)
@
}
\end{frame}

@@

\begin{frame}
\frametitle{Returns}
\framesubtitle{Education returns to behaviour skills in BCS}
\centering
\resizebox{!}{4cm}{%
<<FA_RETURNS, dependson='FA_SCORES_SELECT', results='asis'>>=
# regress scores on income
library(dplyr)
library(stargazer)

@@

\begin{frame}[noframenumbering]
\frametitle{Factor scores vs. Sum scores}
\begin{figure}
<<FACRAW, dev='tikz', fig.height=7, fig.width=7, out.width='.65\\linewidth'>>=
@
\end{figure}
\end{frame}

@@

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Visualising the gradient -- inflation-adjusted income}
\begin{figure}
<<LOESS_INFL_EXT, dependson='LOESS_INFL', dev='tikz', fig.height=8, fig.width=12, out.width='.95\\linewidth'>>=
ploess.infl.ext
@
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Visualising the gradient -- inflation-adjusted income}
\begin{figure}
<<LOESS_INFL_INT, dependson='LOESS_INFL', dev='tikz', fig.height=8, fig.width=12, out.width='.95\\linewidth'>>=
ploess.infl.int
@
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Inequality}
\framesubtitle{Quantifying the gradient -- inflation-adjusted income}
\centering
\resizebox{!}{4cm}{%
<<FA_REGS_INFL, dependson='FA_SCORES_SELECT', results='asis'>>=
# regress scores on income
library(dplyr)
library(stargazer)
